WASP-4 is Accelerating Towards the Earth

### 2020-01-31 paper/ms.tex
more word-smithing

The system's acceleration towards our line-of-sight causes a decrease
in the apparent orbital period.  The expected period derivative is
\begin{align}
  \dot{P}_{\rm\,RV}
  = \frac{\dot{v}_{\rm r} P}{c}
  = -5.94 \pm 0.39~{\rm msec}\,{\rm yr}^{-1}.
\end{align}
In other words, the majority of the period decrease observed in
transits ($\dot{P}= - 8.64 \pm 0.89~{\rm msec}\,{\rm yr}^{-1}$) seems to
be explained by the acceleration of the host star.

@@

\subsection{Implications for WASP-4}
Previous potential explanations for WASP-4b's decreasing orbital
period included tidal orbital decay, orbital precession, and
light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
velocity measurements strongly indicate that the least exotic
option---light-travel time effects---is also the most likely.
Transits show the obital period decreasing by $- 8.64 \pm 0.89~{\rm
ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
perfect, Occam's razor would suggest that the line-of-sight
acceleration is probably a sufficient explanation for the apparent
decrease of WASP-4b's orbital period.

@@

% \citeauthor{baluev_2019} analyzed their newly obtained photometry,
% along with archival light curves that we omitted from our analysis
% due to systematic uncertainties in the absolute time system.
% \citeauthor{baluev_2019} found that when they used all the available
% TTV data, the need for a quadratic ephemeris was present.  However,
% they pointed out that if they used lower-precision subsets of the
% available timing data, the necessity for the quadratic term
% decreased.  This is not particularly surprising.
% \citeauthor{baluev_2019} also pointed out that the precise transit
% times reported by \citet{huitson_gemini_2017} were quite important
% in the time-series.  Overall, \citeauthor{baluev_2019} did not find
% the claim of a period decrease convincing.

@@

Including WASP-4b, 16 of 51 hot Jupiters surveyed by
\citet{knutson_friends_2014} show a non-zero radial velocity trend.
Therefore around 1 in 3 hot Jupiters is expected to show period
changes commensurate with WASP-4 due to acceleration by outer
companions.  Half of these will be period decreases, and will be an
astrophysical false positive in searches for tidal orbital decay.

>>>>

The system's acceleration towards our line-of-sight causes a decrease
in the apparent orbital period:
\begin{align}
  \dot{P}_{\rm\,RV} &= \frac{\dot{v}_{\rm r} P}{c}.
\end{align}
Expressed in convenient units,
\begin{align}
  \dot{P}_{\rm\,RV} &= 105{.}3\,{\rm msec}\,{\rm yr}^{-1} \
  \left( \frac{P}{{\rm day}} \right)
  \left( \frac{\dot{\gamma}}{ {\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}} \right).
\end{align}
For WASP-4, this yields
\begin{align}
  \dot{P}_{\rm\,RV} &= -5.94 \pm 0.39\,{\rm msec}\,{\rm yr}^{-1}.
\end{align}
In other words, the majority of the period decrease observed in
transits ($\dot{P}= - 8.64 \pm 0.89~{\rm msec}\,{\rm yr}^{-1}$) seems to
be explained by the acceleration of the host star.

@@

\subsection{Implications for WASP-4}
Previous potential explanations for WASP-4b's decreasing orbital
period included tidal orbital decay, orbital precession, and
light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
velocity measurements strongly indicate that the least exotic
option---light-travel time effects---is also the most likely.
Transits show the obital period decreasing by $- 8.64 \pm 0.89~{\rm
ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
perfect, Occam's razor would suggest that most or all of the apparent
decrease of WASP-4b's orbital period is caused by the line-of-sight
acceleration.

@@

@@

Including WASP-4b, 16 of 51 hot Jupiters surveyed by
\citet{knutson_friends_2014} show a non-zero radial velocity trend.
Therefore around 1 in 3 hot Jupiters is expected to show period
changes commensurate with WASP-4 due to acceleration by outer
companions.  Half of these will be period decreases\footnote{ At a
population level, one would expect half of the line-of-sight
accelerations to be positive, and half negative.  Tidal orbital decay
however is usually negative, and might manifest as an asymmetry in the
$\dot{P}$ distribution of hot Jupiters.}, and will be an
astrophysical false positive in searches for tidal orbital decay.
